Indirect Conflicts: An Exploration and Discussion of Tools, Process, and Developer Insight

### 2013-08-29 iced-2013/paper.tex
Introduction added large parts.

\author{\IEEEauthorblockN{Jordan Ell}
\IEEEauthorblockA{University of Victoria,
Victoria, Canada \\ jell@uvic.ca}
IEEEauthorblockN{Daniela Damian}
\IEEEauthorblockA{University of Victoria,
Victoria, Canada \\ danielad@csc.uvic.ca}
}

@@

\begin{abstract}
Developer awareness techniques have been proposed and studied to aid in developer
understanding and  efficient developer output in terms of source code produced. These techniques
involve the sharing of information regarding source code changes in a specific project in order
to either catch potential issues before they are fully integrated or to help developers debug
pre-existing issues. These techniques have largely focused on either \textit{direct} or 
\textit{indirect conflict} issues. While the techniques and tools for direct conflicts have had
large success, tools either proposed or studied which focus on indirect conflicts have had many common
struggles and poor developer interest. To better understand why indirect conflict tools either fail or have little
developer investment, we performed a grounded theory study with 97 developers involved in either
interviews or a web survey. We identified the main features that characterize indirect conflict issues,
what processes and techniques developers use to aid in preventing and solving indirect conflict issues,
and what developers would look for in a tool to aid them in these conflicts.

@@

\section{Related Work}

\section{Methodology}

@@

Participants were invited to participate in the interviews by email and were sent a single reminder email one week
after the initial invitation if no response had been made. We directly emailed 22 participants and ended up conducting
19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
interviews were not possible, one of Skype or Google Hangout was used with audio and video being recorded but only
audio being stored for future use. We were quite pleased with the response rate to our initial invitation emails as
our breadth of participants was rather large spanning multiple open and closed source project and our interview 
answer saturation was quite high (to be seen in Section~\ref{sec:rnd}). However, we were disappointed in our word
of mouth campaign. We had asked several participants to pass along our interviews invitations to other members of
their respective organizations but did not receive any additional participants through this method.

@@

Each of the categories was used as a
point of aggregation for later analysis as a way of answering the X research questions laid out in 
Section~\ref{sec:intro}. While each of the 12 question categories had a number of starter questions, interviews 
largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
In general, once the category was explained to the participant, little prompting was necessary as the participant
would jump right into an experience, or issue he or she had with the question or category. However, not all
participants had strong opinions or any experience on every category mentioned. For these participants, answers 
to the specific categories were not required or pressed upon. We attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being apart of any one software project for extended periods of time. We account for these non answers
in our analysis as seen in Section~\ref{sec:rnd}. Interviews lasted 
from 15 minutes up to 75 minutes. The length of the interview largely depended on the amount of experience the
participant had in software development as they were able to give more experiences and more informed opinions.

@@

\section{Results}
\label{sec:rnd}

@@

@@

\section{Discussion}

\section{Future Work}

\section{Conclusions}

@@

%\bibliographystyle{IEEEtran}
%\balance
%\bibliography{paper}

>>>>

\author{\IEEEauthorblockN{Jordan Ell}
\IEEEauthorblockA{University of Victoria,
Victoria, Canada \\ jell@uvic.ca}
\IEEEauthorblockN{Daniela Damian}
\IEEEauthorblockA{University of Victoria,
Victoria, Canada \\ danielad@csc.uvic.ca}
}

@@

\begin{abstract}

Developer awareness techniques have been proposed and studied to aid in developer
understanding and  efficient developer output in terms of source code produced. These techniques
involve the sharing of information regarding source code changes in a specific project in order
to either catch potential issues before they are fully integrated or to help developers debug
pre-existing issues. These techniques have largely focused on either \textit{direct} or 
\textit{indirect conflict} issues. While the techniques and tools for direct conflicts have had
large success, tools either proposed or studied which focus on indirect conflicts have had many common
struggles and poor developer interest. To better understand why indirect conflict tools either fail or have little
developer investment, we performed a grounded theory study with 97 developers involved in either
interviews or a web survey. We identified the main features that characterize indirect conflict issues,
what processes and techniques developers use to aid in preventing and solving indirect conflict issues,
and what developers would look for in a tool to aid them in these conflicts.

@@

As Software Configuration Management (SCM) has grown over the years, the maturity and norm of parallel 
development has become the standard development process instead of the exception. With this parallel development
comes the need for larger awareness among developers to have ``an understanding of the activities of others
which provides a context for one's own activities''~\cite{Dourish:1992:ACS}. This added awareness in theory
mitigates the downside of parallel development which is the cost of conflict prevention and resolution, however,
in practice we see these mitigated losses appear quite frequently. Not only do these conflicts still occur,
but at times they can prove to be a significant and time-consuming chore for developers~\cite{Perry:2001:PCL}.

Two types of conflicts have attracted the attention of academics over the years, \textit{direct} and 
\textit{indirect conflicts}. Direct conflicts involve immediate workspace concerns such as developers editing the same
artifact, finding expert knowledge about a particular file, or knowing when another developer is working on
the same file as you. Tools have been created and studied for these types of issues
~\cite{Xiang:2008:ERT, Biehl:2007:FVD, Sarma:2009:TIV, Khurana:2009:PFC} with relatively good success and 
positive developer feedback. However, indirect conflict tools have not had the same success. Indirect conflicts
can arise in source code for such reasons as having one's own code negatively affected by a library upgrade or
making a code change only to find out it had negative effects on unit or integration test or  having negative effects
on the way other developer's have interpreted the use of your code. Tools that have attempted to either
help developers prevent or solve indirect conflicts have either attracted little developer interest or have
had some common struggles which remain to be studied or solved.

% Another paragraph about indirect conflict tools.

We are interested in exploring three main areas in the hopes of resolving some of the issues that are occurring
with indirect conflict tools: the nature of these conflicts in the how, what, and where, the processes that
developers use prevent, catch, and solve indirect conflicts, and what developer opinions are on preventative 
indirect conflicts tools as well as what they would look for in any tool for indirect conflicts. We broke
these categories down into the following 8 research questions:

\begin{description}
	\item[RQ1] \textit{How do indirect conflicts occur?}
	\item[RQ2] \textit{How often do indirect conflicts occur and at what times in a project do they occur more frequently?}
	\item[RQ3] \textit{What software artifacts and environments are susceptible to indirect conflicts?}
	\item[RQ4] \textit{How do developer prevent indirect conflicts?}
	\item[RQ5] \textit{How do developers catch indirect conflicts?}
	\item[RQ6] \textit{How to developers react to and solve indirect conflicts?}
	\item[RQ7] \textit{When are pre-emptive indirect conflict tools best used and what would developers want from them?}
	\item[RQ8] \textit{What kinds of tools do developers want to help them handle and solve indirect conflicts?}
\end{description}

We interviewed 19 developers from across 12 companies and open source projects as well as surveyed 78 
developers in order to answer these questions.

% Paragraph(s) which summarizes the key result points.

We continue this paper by first presenting our motivations for this research in Section~\ref{sec:mot}.
We then present the summarized methodology we used to perform this study in Section~\ref{sec:meth}. We
continue with the results of our study in regards to our 8 research questions in Section~\ref{sec:results} 
followed by a discussion in Section~\ref{sec:disc} of what was found and how it can be applied to future study and implementation
of indirect tools or processes. We next present how evaluation was performed in terms of the credibility
and quality of our results in Section~\ref{sec:eval}, then cover related work in Section~\ref{sec:related}
and conclude in Section~\ref{sec:conc}.

\section{Motivation}
\label{sec:mot}

\section{Methodology}
\label{sec:meth}

@@

Participants were invited to participate in the interviews by email and were sent a single reminder email one week
after the initial invitation if no response had been made. We directly emailed 22 participants and ended up conducting
19 interviews. Interviews were conducted in person when possible and recorded for audio content only. When in person
interviews were not possible, one of Skype or Google Hangout was used with audio and video being recorded but only
audio being stored for future use. We were quite pleased with the response rate to our initial invitation emails as
our breadth of participants was rather large spanning multiple open and closed source project and our interview 
answer saturation was quite high (to be seen in Section~\ref{sec:results}). However, we were disappointed in our word
of mouth campaign. We had asked several participants to pass along our interviews invitations to other members of
their respective organizations but did not receive any additional participants through this method.

@@

Each of the categories was used as a
point of aggregation for later analysis as a way of answering the X research questions laid out in 
Section~\ref{sec:intro}. While each of the 12 question categories had a number of starter questions, interviews 
largely became discussions of developer experience and opinion as opposed to direct answers to any specific question.
In general, once the category was explained to the participant, little prompting was necessary as the participant
would jump right into an experience, or issue he or she had with the question or category. However, not all
participants had strong opinions or any experience on every category mentioned. For these participants, answers 
to the specific categories were not required or pressed upon. We attribute any non answer by a participant to
either lack of knowledge in their current project pertaining to the category or lack of experience in terms of
being apart of any one software project for extended periods of time. We account for these non answers
in our analysis as seen in Section~\ref{sec:results}. Interviews lasted 
from 15 minutes up to 75 minutes. The length of the interview largely depended on the amount of experience the
participant had in software development as they were able to give more experiences and more informed opinions.

@@

\section{Results}
\label{sec:results}

@@

\begin{description}
	\item[RQ4] \textit{How do developer prevent indirect conflicts?}
\end{description}

% Design by contract, add and deprecate

@@

\section{Discussion}
\label{sec:disc}

\section{Evaluation}
\label{sec:eval}

\section{Related Work}
\label{sec:related}

\section{Conclusions}
\label{sec:conc}

@@

\bibliographystyle{IEEEtran}
\balance
\bibliography{paper}
